---
title: "Cleaning HMO Data"
---

```{r}
library(tidyverse)
```


```{r}
hmo = read_csv(url("https://intro-datascience.s3.us-east-2.amazonaws.com/HMO_data.csv"))
```
It appears that the "X" column is functionally useless, so let's remove it for now
```{r}
hmo = hmo %>% dplyr::select(-X)
```

Many of the variables are categorical with two values. Let's make as descriptive
as possible.

```{r}
hmo = hmo %>% 
  mutate(smoker = factor(ifelse(smoker == "yes", "smoker", "nonsmoker")),
         hypertension = factor(ifelse(hypertension > 0, "yes", "no")))
```

Let's also turn all other categorical variables into factors

```{r}
hmo = hmo %>% mutate(location        = as.factor(location),
                     location_type   = as.factor(location_type),
                     education_level = as.factor(education_level),
                     yearly_physical = as.factor(yearly_physical),
                     exercise        = as.factor(exercise),
                     married         = as.factor(married),
                     gender          = as.factor(gender))
```

Expensive is defined as having cost greater than 5000
```{r}
cost_threshold = 5000
hmo = hmo %>% mutate(expensive = factor(ifelse(cost > cost_threshold, "yes", "no")))
```

At this point, let's save this as a csv

```{r}
saveRDS(hmo, "hmo_cleaned_with_numeric.RData")
```

Let's also create a version that turns age, bmi, and children into categorical
variables, while eliminating cost

```{r}
bmi_missing_id = which(is.na(hmo$bmi))
hmo$bmi[bmi_missing_id] = 0
hmo = hmo %>%
  mutate(age = cut(age,
                   breaks = c(0,20,30,40,50,65,200),
                   labels = c("<20", "20-29", "30-39","40-49", "50-65", ">65"), right = F),
         bmi = cut(bmi,
                   breaks = c(0, 18.5, 25, 30, 35, 100),
                   labels = c("underweight", "normal", "overweight", "obese", "extremely obese"),right = F,),
         children = as.factor(children)) %>%
  select(-c(cost))
hmo$bmi[bmi_missing_id] = NA
```

Let's write this version as a separate csv

```{r}
saveRDS(hmo, "hmo_cleaned_all_factors.RData")

```













